The combination of reporting bias and underpowered study designs has substantially exaggerated the motor learning benefits of self-controlled practice and enhanced expectancies: A meta-analysis

### 2022-09-28 docs/submission/manuscript.tex
update: small fixes to match journal submission instructions

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  doc, donotrepeattitle,floatsintext]{apa7}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\ifLuaTeX
\usepackage[bidi=basic]{babel}
\else
\usepackage[bidi=default]{babel}
\fi
\babelprovide[main,import]{english}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

@@

\usepackage{xpatch}
\makeatletter
\xapptocmd\appendix
  {\xapptocmd\section
    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
    {}{\InnerPatchFailed}%
  }
{}{\PatchFailed}
\keywords{Motor learning, OPTIMAL theory, Publication bias, Z-curve}
\usepackage{csquotes}
\usepackage{pdflscape}
\usepackage{longtable}
\usepackage{graphicx}
\usepackage{graphbox}
\usepackage{tabularray}
\UseTblrLibrary{booktabs}
\raggedbottom
\renewcommand\author[1]{}
\renewcommand\affiliation[1]{}
\authorsnames[1, 2, 3, 3, 1]{Brad McKay, Mariane F.B. Bacelar, Juliana O. Parma, Matthew W. Miller\textsuperscript{$\dagger$}, Michael J. Carter\textsuperscript{$\dagger$}\vspace{2ex}}
\authorsaffiliations{{Department of Kinesiology, McMaster University}, {Department of Kinesiology, Boise State University}, {School of Kinesiology, Auburn University}}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={The combination of reporting bias and underpowered study designs has substantially exaggerated the motor learning benefits of self-controlled practice and enhanced expectancies: A meta-analysis},
  pdflang={en-EN},
  pdfkeywords={Motor learning, OPTIMAL theory, Publication bias, Z-curve},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

@@

\abstract{%
Enhanced expectancies and autonomy-support through self-controlled practice conditions form the motivation pillar of OPTIMAL theory (Wulf \& Lewthwaite, 2016). The influence of these practice variables on motor learning was recently evaluated in two separate meta-analyses. Both meta-analyses found that the published literature suggested a moderate and significant benefit on motor learning; however, evidence for reporting bias was found in both literatures. Although multiple bias-corrected estimates were reported in the self-controlled meta-analysis, there was no principled way to prefer one over the other. In the enhanced expectancies meta-analysis, the trim-and-fill-technique failed to correct the estimated effects. Here, we addressed these limitations by reanalyzing the data from both meta-analyses using robust Bayesian meta-analysis methods. Our reanalysis revealed that reporting bias substantially exaggerated the benefits of these practice variables in the original meta-analyses. The true effects are instead small, uncertain, and potentially null. We also found the estimated average statistical power among all studies from the original meta-analyses was 6\% (95\% confidence interval {[}5\%, 13\%{]}). These results provide compelling and converging evidence that strongly suggests the available literature is insufficient to support the motivation pillar of OPTIMAL theory. Our results highlight the need for adequately powered experimental designs if motor learning scientists want to make evidence-based recommendations.
}

@@

Addressing reporting bias presents substantial challenges to meta-analysts. Since reporting bias limits the information we have access to, it is impossible to know for certain how much bias is present or how large the impact is (Carter et al., 2019; McShane et al., 2016). The best we can do is think carefully about the mechanisms that potentially underlie reporting bias and attempt to model them accurately. In contrast, a naive random effects analysis assumes there is no reporting bias. Complicating matters, reporting bias can take several different forms with unique impacts on the final sample (Maier et al., in-press; Stefan \& Schönbrodt, 2022; Thornton \& Lee, 2000). To account for this, multiple models of reporting bias need to be attempted without knowing which is most likely \emph{a priori}. Thus, bias-correction analyses are inherently sensitivity analyses (Mathur \& VanderWeele, 2020; Sutton et al., 2000; Vevea \& Woods, 2005). Until recently, if the results of multiple sensitivity analyses differed widely, there was no mechanism to reconcile the estimates. We now have Robust Bayesian Meta-Analysis (RoBMA) methods that apply Bayesian model averaging to allow meta-analysts to fit several plausible models (see Table \ref{tab:table1} for descriptions) and give greater weight to the models that best account for the data (Bartoš et al., 2022; Maier et al., in-press). The RoBMA method provides single estimates of the average effect and heterogeneity, along with Bayes factors to quantify the evidence in support of a true effect, the presence of heterogeneity, and the presence of reporting bias. Results from simulation studies and analyses of real data with known reporting bias mechanisms suggest that RoBMA is substantially more accurate and less biased than naive random effects models and also performs better than other competing reporting bias models (Bartoš et al., 2022).

\clearpage

\renewcommand{\arraystretch}{1.7}
\DefTblrTemplate{caption-tag}{default}{Table\hspace{0.25em}\thetable}
\DefTblrTemplate{caption-sep}{default}{.\enskip}
\DefTblrTemplate{caption-text}{default}{\InsertTblrText{caption}}
\DefTblrTemplate{contfoot-text}{default}{(\emph{Continued})}
\DefTblrTemplate{middlehead,lasthead}{default}{\textbf{Table 1.} Continued}
\SetTblrStyle{caption-tag,caption-sep}{\bfseries}
\small
\begin{longtblr}[
  caption = {The selection and regression models used in our robust Bayesian meta-analysis approach.},
  label = {tab:table1},
]{colspec = {p{4.5cm}p{5cm}p{4.5cm}}}
\toprule
  \textbf{Type of selection} &
  \textbf{Visualization} &
  \textbf{Example scenario} \\ 
\midrule
  \textbf{\underline{Selection models}} &
  &
  \textbf{\underline{Direction not important}} \\
Significant results are more likely to be reported in either direction (two-tailed)  &
  \raisebox{-0.9\height}{\includegraphics[width=4.75cm]{../../figs/model1.png}} &
  Researcher conducts test and observes a null result. They decide the experiment did not work and move on. Significant results get reported. \\
Significant results are most likely to be reported, but ``non-significant trends'' are more likely to be reported than other null results in either direction (two-tailed). &
  \raisebox{-0.9\height}{\includegraphics[width=4.75cm]{../../figs/model2.png}} &
  Authors report significant results and ``non-significant trends''. The latter may be interpreted as fair evidence the manipulation worked. Some reviewers take issue with trends, so only some make it through and get reported. Null results unlikely to be written up. \vspace{1em} \\
  &
  &
  \textbf{\underline{Direction important}} \\
Significant results and non-significant trends are more likely to be reported in the predicted direction. &
  \raisebox{-0.9\height}{\includegraphics[width=4.75cm]{../../figs/model3.png}} &
  Researcher is confident in the hypothesis being tested in an experiment and doubts the validity of null or opposing findings. Reports results they are confident in. \\
Significant results in the predicted direction are more likely to be reported than trends, which are more likely to be reported than other null results and significant results in the opposite (i.e., ``wrong'') direction. &
  \raisebox{-0.9\height}{\includegraphics[width=4.75cm]{../../figs/model4.png}} &
  A preference for reporting findings with a compelling narrative results in preferring significant results and occasionally trends. Null or conflicting results less likely to add to the narrative. \\
Significant results and trends in the predicted (i.e., ``correct'') direction are more likely to be reported than null findings in the predicted (i.e., ``correct'') direction, which are more likely to be reported than results in the opposite (i.e., ``wrong'') direction. &
  \raisebox{-0.9\height}{\includegraphics[width=4.75cm]{../../figs/model5.png}} &
  A student observes results in the opposite direction of what was expected. Supervisor thinks something may have went wrong so results not published. Other students publish results consistent with predictions. \\
Full selection model. Significant results most likely, then trends, then null results in the predicted (i.e., ``correct'') direction. The least likely to be reported are results in the opposite (i.e., ``wrong'') direction. &
  \raisebox{-0.9\height}{\includegraphics[width=4.75cm]{../../figs/model6.png}} &
  An editor prefers to publish interesting results. Prediction successes are interesting. Some trends are interesting if they are believable. Results in the opposite direction are interesting, but only if replicated. \\
  \textbf{\underline{Regression models}} &
  &
  \\
Conditioning on smaller \emph{p}-values in the predicted direction creates a relationship between effect sizes and standard errors. Called ``small study effects'' because all else being equal smaller studies need larger effects to achieve significant results. &
  \raisebox{-0.9\height}{\includegraphics[width=4.75cm]{../../figs/model7.png}} &
  This models the dependency caused by selective reporting, not the underlying mechanism itself. Dependency can be caused by a third variable, such as intensity of the interventions used in smaller compared to larger studies. \\
Quadratic relationship between effect and standard errors. Large studies likely to be reported independent of results, while smaller studies need increasingly large effects in the predicted (i.e., ``correct'') direction to avoid censorship. &
  \raisebox{-0.9\height}{\includegraphics[width=4.75cm]{../../figs/model8.png}} &
  Researchers invest in conducting a large study and are motivated to publish regardless of the results. They persevere if null results are rejected. Small studies are abandoned unless the results are impressive. \\
\bottomrule
\end{longtblr}

\clearpage

\normalsize

@@

Data and code used in this study can be accessed using either of the following links: \url{https://osf.io/vfza7/} or \url{https://github.com/cartermaclab/proj_sc-ee-optimal-theory}.

@@

The original meta-analyses followed similar inclusion criteria and data extraction policies (see Figure \ref{fig:fig1}). All data were extracted by two independent researchers with conflicts settled by including a third researcher. The literature search for the self-controlled practice meta-analysis (for further details see McKay et al., in-press) ended in August 2019 and in June 2020 for the enhanced expectancies meta-analysis (for further details see Bacelar, Parma, Murrah, et al., 2022).

\clearpage

\begin{figure}

{\centering \includegraphics{../../figs/fig1}

\caption{\normalfont \footnotesize \textbf{Description of the original meta-analyses.} Diagram outlining the shared and unique inclusion criteria, dependent variable selection, the number of studies screened, and the number of studies included in each meta-analysis.}\label{fig:fig1}
\end{figure}

\clearpage

\normalsize

@@

The second class of models fit in the RoBMA-PSMA method model the relationship between observed effect sizes and their standard errors. A positive relationship between effect sizes and their standard errors has been termed ``small study effects'' (Sterne et al., 2000). The RoBMA-PSMA includes the precision-effect test (PET) and the precision-effect test with standard errors (PEESE) for small study effects. The PET model fits the relationship between observed effect sizes and their standard errors, while the PEESE model instead includes the square of the standard errors (i.e., their variances) (Stanley \& Doucouliagos, 2014). The difference between the PET and PEESE models is that the PET model fits a linear relationship between the effect observed in a study and the precision with which that effect was estimated. The PEESE model fits a quadratic relationship. Thus, the PET and PEESE models differ with respect to the assumed underlying selection process (Stanley \& Doucouliagos, 2014). The PET model assumes that effect sizes increase in step with decreases in precision, consistent with selection based on statistical significance. The PEESE model assumes that studies with high precision are likely to be published regardless of statistical significance, whereas increasingly imprecise studies require increasingly larger effect size estimates to survive censorship.

@@

\hypertarget{z-curve}{%
\subsection{\texorpdfstring{\emph{Z}-curve}{Z-curve}}\label{z-curve}}

@@

The model-averaged posterior distribution of the average effect from the RoBMA-PSMA model is displayed in Figure \ref{fig:fig2}B. The results suggest moderate evidence against the presence of an effect, BF\textsubscript{01} = 3.16, very weak evidence against the presence of heterogeneity, BF\textsubscript{rf} = 1.7, and overwhelming evidence for the presence of reporting bias, BF\textsubscript{pb} = 18,399. The overall model ensemble estimated the effect of self-controlled practice as \emph{d} = .034 (95\% credible interval {[}.0, .248{]}). Heterogeneity was estimated as \(\tau\) = .05 (95\% credible interval {[}.0, .261{]}). A model fit with two influential cases included found overwhelming evidence for the presence of heterogeneity, BF\textsubscript{rf} = 1,924,516 and estimated \(\tau\) = .559 (95\% credible interval {[}.36, .78{]}). There were no other meaningful differences between models.

@@

The model-averaged posterior distribution of the average effect from the RoBMA-PSMA model is displayed in Figure \ref{fig:fig2}C. The results revealed weak evidence for the presence of an effect BF\textsubscript{10} = 1.9, weak evidence for the presence of reporting bias, BF\textsubscript{pb} = 2.3, and very strong evidence for the presence of heterogeneity, BF\textsubscript{rf} = 47.6. The model ensemble estimated an average effect of \emph{d} = .26 (95\% credible interval {[}-.07, .63{]}). Heterogeneity was estimated as \(\tau\) = .35 (95\% credible interval {[}.07, .54{]}). A model fit with two influential cases included found weak evidence for the absence of an effect BF\textsubscript{01} = 2.5, strong evidence for the presence of reporting bias, BF\textsubscript{pb} = 21, and overwhelming evidence for the presence of heterogeneity, BF\textsubscript{rf} = 45,300. The model ensemble estimated an average effect of \emph{d} = .00 (95\% credible interval {[}-.62, .59{]}), \(\tau\) = .49 (95\% credible interval {[}.32, .68{]}).

@@

\clearpage

\begin{figure}

{\centering \includegraphics{../../figs/fig2}

\caption{\normalfont \footnotesize \textbf{Results from the robust Bayesian meta-analysis with publication selection model-averaging (RoBMA-PSMA) method.} \textbf{(A)} Prior spike and distribution (purple) with 50\% of prior probability density concentrated on the null hypothesis and 50\% reflecting plausible true effects with \emph{M} = 0 and \emph{SD} = 1. (The x-axis is truncated; the prior did include a small probability of effects larger than 1 and smaller than -1.) The model-averaged posterior distribution of the mean effect (Mu) of \textbf{(B)} self-controlled practice on motor learning (blue). The increased height of the spike at mu = 0 reflects increased belief in the null hypothesis. The remaining distribution reflects updated belief in the size and direction of a possible true effect. The mean estimate (\emph{M} = .034) for the posterior distribution is represented by the dashed vertical line. The model-averaged posterior distribution of the mean effect of \textbf{(C)} enhanced expectancies on motor learning (red). The decreased height of the spike at mu = 0 reflects decreased belief in the null hypothesis. The remaining distribution reflected updated belief in the size and direction of a possible true effect. The mean estimate (\emph{M} = .26) for the posterior distribution is represented by the dashed vertical line. The model-averaged posterior distribution generated from an analysis of \textbf{(D)} simulated data (green). In the simulation, 49 studies were sampled from a population with a true effect of mu = .54 and no reporting bias. The mean estimate (\emph{M} = .504) for the posterior is represented by the vertical dashed line. Outliers were not included in the self-controlled (\emph{n} = 2) and enhanced expectancies (\emph{n} = 2) models.}\label{fig:fig2}
\end{figure}

\clearpage

\normalsize

\hypertarget{z-curve-1}{%
\subsection{\texorpdfstring{\emph{Z}-curve}{Z-curve}}\label{z-curve-1}}

@@

The results of the \emph{z}-curve analysis can be seen in Figure \ref{fig:fig3}B. The analysis estimated the average statistical power (expected discovery rate) of all experiments examining the effect of self-controlled practice was 6\% (95\% confidence intervals {[}5\%, 12\%{]}). Since the upper bound of the expected discovery rate does not overlap with the observed discovery rate of 48\% (95\% confidence interval {[}35\%, 62\%{]}), there is evidence of significant reporting bias. The estimated conditional power of the statistically significant results (expected replication rate) was 11\% (95\% confidence interval {[}3\%, 30\%{]}). Including influential cases did not markedly change the results.

@@

The results of the \emph{z}-curve analysis can be seen in Figure \ref{fig:fig3}C. The analysis estimated that the expected discovery rate of studies conducted on enhanced expectancies was 8\% (95\% confidence interval {[}5\%, 18\%{]}). The observed discovery rate was 44\% (95\% confidence interval {[}31\%, 59\%{]}). Since the upper bound of the expected discovery rate does not overlap with the lower bound of the observed discovery rate, there is evidence of significant reporting bias. The expected replication rate of the statistically significant results was 33\% (95\% confidence interval {[}8\%, 59\%{]}). Including influential cases did not meaningfully change the results.

@@

The results of the \emph{z}-curve analysis of enhanced expectancy and self-controlled practice meta-data combined can be seen in Figure \ref{fig:fig3}D. The expected discovery rate of all studies conducted on the motivational factors in OPTIMAL theory is 6\% (95\% confidence interval {[}5\%, 13\%{]}). The observed discovery rate was 46\% (95\% confidence interval {[}37\%, 56\%{]}). The lower bound of the observed discovery rate does not overlap with the upper bound of the expected discovery rate, providing evidence of significant reporting bias in this literature. The expected replication rate of the statistically significant results was 21\% (95\% confidence interval {[}4\%, 39\%{]}). Including influential cases did not meaningfully change the results.

\clearpage

\begin{figure}

{\centering \includegraphics{../../figs/fig3}

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  man, donotrepeattitle,mask,floatsintext]{apa7}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\ifLuaTeX
\usepackage[bidi=basic]{babel}
\else
\usepackage[bidi=default]{babel}
\fi
\babelprovide[main,import]{english}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

@@

\usepackage{xpatch}
\makeatletter
\xapptocmd\appendix
  {\xapptocmd\section
    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
    {}{\InnerPatchFailed}%
  }
{}{\PatchFailed}
\keywords{OPTIMAL theory, Motivation, RoBMA, \emph{Z}-curve, Publication bias\newline\indent Word count: 6354}
\usepackage{lineno}

\linenumbers
\usepackage{csquotes}
\usepackage{newtxtext,newtxmath}
\usepackage{pdflscape}
\usepackage{longtable}
\usepackage{graphicx}
\usepackage{graphbox}
\usepackage{tabularray}
\UseTblrLibrary{booktabs}
\raggedbottom
\usepackage{sectsty}
\sectionfont{\centering\normalfont\normalsize\bfseries}
\subsectionfont{\normalfont\normalsize\bfseries\itshape}
\subsubsectionfont{\normalfont\normalsize\itshape}
\usepackage{censor}
\renewcommand\author[1]{}
\renewcommand\affiliation[1]{}
\authorsnames[1, 2, 3, 3, 1]{Brad McKay, Mariane F.B. Bacelar, Juliana O. Parma, Matthew W. Miller\textsuperscript{$\dagger$}, Michael J. Carter\textsuperscript{$\dagger$}\vspace{2ex}}
\authorsaffiliations{{Department of Kinesiology, McMaster University}, {Department of Kinesiology, Boise State University}, {School of Kinesiology, Auburn University}}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={The combination of reporting bias and underpowered study designs has substantially exaggerated the motor learning benefits of self-controlled practice and enhanced expectancies: A meta-analysis},
  pdflang={en-EN},
  pdfkeywords={OPTIMAL theory, Motivation, RoBMA, \emph{Z}-curve, Publication bias},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

@@

\abstract{%
Enhanced expectancies and autonomy-support through self-controlled practice conditions form the motivation pillar of OPTIMAL theory (Wulf \& Lewthwaite, 2016). The influence of these practice variables on motor learning was recently evaluated in two separate meta-analyses. Both meta-analyses found that the published literature suggested a moderate and significant benefit on motor learning; however, evidence for reporting bias was found in both literatures. Although multiple bias-corrected estimates were reported in the self-controlled meta-analysis, there was no principled way to prefer one over the other. In the enhanced expectancies meta-analysis, the trim-and-fill-technique failed to correct the estimated effects. Here, we addressed these limitations by reanalyzing the data from both meta-analyses using robust Bayesian meta-analysis methods. Our reanalysis revealed that reporting bias substantially exaggerated the benefits of these practice variables in the original meta-analyses. The true effects appear small, uncertain, and potentially null. We found the estimated average statistical power among all studies from the original meta-analyses was 6\% (95\% confidence interval {[}5\%, 13\%{]}). These results provide compelling and converging evidence that strongly suggests the available literature is insufficient to support the motivation pillar of OPTIMAL theory. Our results highlight the need for adequately powered experimental designs if motor learning scientists want to make evidence-based recommendations.
}

@@

Addressing reporting bias presents substantial challenges to meta-analysts. Since reporting bias limits the information we have access to, it is impossible to know for certain how much bias is present or how large the impact is (Carter et al., 2019; McShane et al., 2016). The best we can do is think carefully about the mechanisms that potentially underlie reporting bias and attempt to model them accurately. In contrast, a naive random effects analysis assumes there is no reporting bias. Complicating matters, reporting bias can take several different forms with unique impacts on the final sample (Maier et al., in-press; Stefan \& Schönbrodt, 2022; Thornton \& Lee, 2000). To account for this, multiple models of reporting bias need to be attempted without knowing which is most likely \emph{a priori}. Thus, bias-correction analyses are inherently sensitivity analyses (Mathur \& VanderWeele, 2020; Sutton et al., 2000; Vevea \& Woods, 2005). Until recently, if the results of multiple sensitivity analyses differed widely, there was no mechanism to reconcile the estimates. We now have Robust Bayesian Meta-Analysis (RoBMA) methods that apply Bayesian model averaging to allow meta-analysts to fit several plausible models (see Table 1 for descriptions) and give greater weight to the models that best account for the data (Bartoš et al., 2022; Maier et al., in-press). The RoBMA method provides single estimates of the average effect and heterogeneity, along with Bayes factors to quantify the evidence in support of a true effect, the presence of heterogeneity, and the presence of reporting bias. Results from simulation studies and analyses of real data with known reporting bias mechanisms suggest that RoBMA is substantially more accurate and less biased than naive random effects models and also performs better than other competing reporting bias models (Bartoš et al., 2022).

\begin{center}
[TABLE 1 NEAR HERE]
\end{center}

@@

Data and code used in this study can be accessed using either of the following links: \url{https://osf.io/vfza7/} or \censor{$https://github.com/cartermaclab/proj_sc-ee-optimal-theory$}.

@@

The original meta-analyses followed similar inclusion criteria and data extraction policies (see Figure 1). All data were extracted by two independent researchers with conflicts settled by including a third researcher. The literature search for the self-controlled practice meta-analysis (for further details see McKay et al., in-press) ended in August 2019 and in June 2020 for the enhanced expectancies meta-analysis (for further details see Bacelar, Parma, Murrah, et al., 2022).

\begin{center}
[FIGURE 1 NEAR HERE]
\end{center}

@@

The second class of models fit in the RoBMA-PSMA method model the relationship between observed effect sizes and their standard errors. A positive relationship between effect sizes and their standard errors has been termed `small study effects' (Sterne et al., 2000). The RoBMA-PSMA includes the precision-effect test (PET) and the precision-effect test with standard errors (PEESE) for small study effects. The PET model fits the relationship between observed effect sizes and their standard errors, while the PEESE model instead includes the square of the standard errors (i.e., their variances) (Stanley \& Doucouliagos, 2014). The difference between the PET and PEESE models is that the PET model fits a linear relationship between the effect observed in a study and the precision with which that effect was estimated. The PEESE model fits a quadratic relationship. Thus, the PET and PEESE models differ with respect to the assumed underlying selection process (Stanley \& Doucouliagos, 2014). The PET model assumes that effect sizes increase in step with decreases in precision, consistent with selection based on statistical significance. The PEESE model assumes that studies with high precision are likely to be published regardless of statistical significance, whereas increasingly imprecise studies require increasingly larger effect size estimates to survive censorship.

@@

\hypertarget{z-curve}{%
\subsection{Z-curve}\label{z-curve}}

@@

The model-averaged posterior distribution of the average effect from the RoBMA-PSMA model is displayed in Figure 2B. The results suggest moderate evidence against the presence of an effect, BF\textsubscript{01} = 3.16, very weak evidence against the presence of heterogeneity, BF\textsubscript{rf} = 1.7, and overwhelming evidence for the presence of reporting bias, BF\textsubscript{pb} = 18,399. The overall model ensemble estimated the effect of self-controlled practice as \emph{d} = .034 (95\% credible interval {[}.0, .248{]}). Heterogeneity was estimated as \(\tau\) = .05 (95\% credible interval {[}.0, .261{]}). A model fit with two influential cases included found overwhelming evidence for the presence of heterogeneity, BF\textsubscript{rf} = 1,924,516 and estimated \(\tau\) = .559 (95\% credible interval {[}.36, .78{]}). There were no other meaningful differences between models.

@@

The model-averaged posterior distribution of the average effect from the RoBMA-PSMA model is displayed in Figure 2C. The results revealed weak evidence for the presence of an effect BF\textsubscript{10} = 1.9, weak evidence for the presence of reporting bias, BF\textsubscript{pb} = 2.3, and very strong evidence for the presence of heterogeneity, BF\textsubscript{rf} = 47.6. The model ensemble estimated an average effect of \emph{d} = .26 (95\% credible interval {[}-.07, .63{]}). Heterogeneity was estimated as \(\tau\) = .35 (95\% credible interval {[}.07, .54{]}). A model fit with two influential cases included found weak evidence for the absence of an effect BF\textsubscript{01} = 2.5, strong evidence for the presence of reporting bias, BF\textsubscript{pb} = 21, and overwhelming evidence for the presence of heterogeneity, BF\textsubscript{rf} = 45,300. The model ensemble estimated an average effect of \emph{d} = .00 (95\% credible interval {[}-.62, .59{]}), \(\tau\) = .49 (95\% credible interval {[}.32, .68{]}).

@@

\begin{center}
[FIGURE 2 NEAR HERE]
\end{center}

\hypertarget{z-curve-1}{%
\subsection{Z-curve}\label{z-curve-1}}

@@

The results of the \emph{z}-curve analysis can be seen in Figure 3B. The analysis estimated the average statistical power (expected discovery rate) of all experiments examining the effect of self-controlled practice was 6\% (95\% confidence intervals {[}5\%, 12\%{]}). Since the upper bound of the expected discovery rate does not overlap with the observed discovery rate of 48\% (95\% confidence interval {[}35\%, 62\%{]}), there is evidence of significant reporting bias. The estimated conditional power of the statistically significant results (expected replication rate) was 11\% (95\% confidence interval {[}3\%, 30\%{]}). Including influential cases did not markedly change the results.

@@

The results of the \emph{z}-curve analysis can be seen in Figure 3C. The analysis estimated that the expected discovery rate of studies conducted on enhanced expectancies was 8\% (95\% confidence interval {[}5\%, 18\%{]}). The observed discovery rate was 44\% (95\% confidence interval {[}31\%, 59\%{]}). Since the upper bound of the expected discovery rate does not overlap with the lower bound of the observed discovery rate, there is evidence of significant reporting bias. The expected replication rate of the statistically significant results was 33\% (95\% confidence interval {[}8\%, 59\%{]}). Including influential cases did not meaningfully change the results.

@@

The results of the \emph{z}-curve analysis of enhanced expectancy and self-controlled practice meta-data combined can be seen in Figure 3D. The expected discovery rate of all studies conducted on the motivational factors in OPTIMAL theory is 6\% (95\% confidence interval {[}5\%, 13\%{]}). The observed discovery rate was 46\% (95\% confidence interval {[}37\%, 56\%{]}). The lower bound of the observed discovery rate does not overlap with the upper bound of the expected discovery rate, providing evidence of significant reporting bias in this literature. The expected replication rate of the statistically significant results was 21\% (95\% confidence interval {[}4\%, 39\%{]}). Including influential cases did not meaningfully change the results.

\begin{center}
[FIGURE 3 NEAR HERE]
\end{center}
